Bootstrapping Multi-atlas Hippocampal Segmentation with \mb

### 2013-04-05 paper.Rnw
Updates to figures with captions and sensible sizes, fixed latex references to figures

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Sweave/Knitr init
<<setup, include=FALSE, cache=FALSE>>=
#opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
#    fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE)
library(ggplot2)
library(Hmisc)
library(reshape2)
library(ADNIMERGE)
library(plyr)
options(digits=3)
theme_set(theme_bw())

@@

\newcommand{\mb}{MAGeT Brain}
\newcommand{\ants}{ANTS}
\newcommand{\animal}{ANIMAL}

@@

\subsubsection{ADNI1 1.5T Complete 1Yr Dataset}
Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
standardized dataset available from ADNI \footnote{
\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
\citep{Wyman2012}. This image collection contains uniformly preprocessed images
which have been designated to be the "best" after quality control.  All images
were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
Systems or Siemens Medical Solutions) at multiple sites using the protocol
described in \citep{Jack2008a}.  Representative 1.5T imaging parameters were TR
= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
x 240mm, a 192 x 192 x 166 matrix (x, y, and z directions) yielding a voxel
resolution of 1.25 x 1.25 x 1.2 $mm^3$. Clinical and demographic data are shown
in table \ref{tab:ADNI1_1.5T_Complete 1Yr_Dataset_Demographics}.

@@

latex(tab, file="", size="scriptsize", 
    caption="ADNI1 1.5T_Complete_1Yr dataset demographics",
    title="tab:ADN1_1.5T_Complete_1Yr_Dataset_Demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Scanning was carried out at the Montreal Neurological Institute on a 1.5-T
Siemens whole body MRI system.  Structural T1 volumes were acquired for each
participant using a three-dimensional (3D) gradient echo pulse sequence with
sagittal volume excitation (repetition time=22ms, echo time=9.2ms, flip
angle=$30{\circ}$, 180 1mm contiguous sagittal slices). The rectangular field-of-view
for the images was 256mm (SI)×204mm (AP).  Subject demographics are shown in
table \ref{tab:SZFEP_Demographics}.

@@

\begin{figure}
<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
               data=fep_sz_demographics, method = "reverse", test = FALSE)
latex(tab, file="", size="scriptsize", ctable=FALSE,
      caption="Schizophrenia First Episode Patient Demographics",
      title="tab:SZFEP_Demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \caption{
  }
\end{figure}

@@

The \animal algorithm carries out image registration is in two phases. In the
first, a 12-parameter linear transformation (3 translations, rotations, scales,
shears) is estimated between images using an algorithm that maximizes the
correlation between blurred MR intensities and gradient magnitude over the whole
brain \citep{Collins}. In the second phase, nonlinear registration is completed
using the \animal algorithm \citep{Collins1995}: an iterative procedure that
estimates a 3D deformation field between two MR images. At first, large
deformations are estimated using blurred version of the input data. These larger
deformations are then input to subsequent steps where the fit is refined by
estimating smaller deformations on data blurred with a Gaussian kernel with a
smaller FWHM. The final transformation is a set of local translations defined on
a bed of equally spaced nodes that were estimated through the optimization of
the correlation coefficient.  For the purposes of this work we used the
regularization parameters optimized in Robbins et al. \citep{Robbins2004},
displayed in table \ref{tab:ANIMAL_params}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ANIMAL parameters                                                           %
%                                                                             %
\begin{table*}[!tbp]
\scriptsize
\caption{\animal Registration Parameters\label{tab:ANIMAL_params}} 
\begin{center}
\begin{tabular}{l | c c c}
\hline 
Parameters        & Stage 1 & Stage 2 & Stage 3  \tabularnewline
\hline 
Model Blur (FWHM) &   8     &    8    &   4      \tabularnewline
Input Blur (FWHM) &   8     &    8    &   4      \tabularnewline
Iterations        &   30    &    30   &   10     \tabularnewline
Step              &   8x8x8 &  4x4x4  &   2x2x2  \tabularnewline
Sub-Lattice       &   6     &    6    &   6      \tabularnewline
Lattice Diameter  &24x24x24 &12x12x12 & 6x6x6    \tabularnewline
\hline
\end{tabular}
\end{center}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

{\bf Dataset evaluated.} 69 1.5T images were arbitrarily selected from the {\em
ADNI1:Complete 1Yr 1.5T} standardized dataset. 23 subjects were chosen from each
disease category (cognitively normal (CN), MCI and AD). Demographics for this
subset are shown in Table \ref{tab:ADNI1_xvalidation_demographics}.

@@

latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption="ADNI-1 cross-validation subset demographics", 
      title="tab:ADNI1_xvalidation_demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

{\bf Atlas and template libary.} The Winterburn atlases were used as the atlas
library.  The template library consisted of 21 randomly selected images from the
ADNI1 data dataset (7 healthy, MCI and AD subjects). Clinical and demographic
data for the template library subjects are shown in Table
\ref{tab:ADNI1_scr_tmpllib_demographics}.

@@

latex(tab, landscape=FALSE, ctable=FALSE, file="", size="scriptsize",
      caption="ADNI1 Complete 1Yr Template Library demographics", 
      title="tab:ADNI1_scr_tmpllib_demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}
<<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=

ggplot(all_data_mean, 
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + 
  facet_grid(reg_method~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1)) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean DSC" )
@
  \label{fig:ADNI1-xval-k_mean}
  \caption{
  {\em Mean DSC by atlas and template library size, registration algorithm,
  and label fusion method. The mean DSC score is calculated over all subjects 
  from the ADNI1 Cross-Validation dataset for each parameter setting.} 
  }
\end{figure}

@@

\begin{figure}
<<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ggplot(subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &
                templates.mb %in% seq(1,20,2)), 
       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
  #facet_grid(reg_method~method.y) + 
  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
  stat_summary(fun.y=mean, geom="point", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  stat_summary(fun.y=mean, geom="line", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  #stat_smooth(method=lm, formula=y~log(x)) + 
  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
  coord_cartesian(ylim=c(-0.1,0.05)) + 
  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(legend=FALSE) + 
  scale_linetype_discrete(legend=FALSE) + 
  xlab( "Number of Templates" ) + 
  ylab( "Increase in mean DSC between MAGeT brain and multiatlas" ) 
@
  \caption{The difference in mean DSC between \mb and multi-atlas segmentations for 
  a range of parameter settings.}
  \label{ADNI1-xval-k-diff}
\end{figure}

@@

ggplot(stats.tests, 
  aes(x=templates.mb, y=p.value, colour=as.factor(atlases), 
      size = abs(statistic))) + 
  facet_grid( . ~ stat) + 
  geom_line(aes(size=0.1)) + 
  geom_point(aes(shape = factor(direction))) + 
  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(guide='none') + 
  scale_linetype_discrete(guide='none') + 
  #scale_size_continuous(guide='none') + 
  scale_shape_discrete("t-statistic sign", labels=c("positive","negative")) + 
  scale_size_area(name="t-statistic magnitude") + 
  scale_x_continuous(breaks=seq(1,20,by=2)) + 
  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
  xlab( "Number of Templates" ) + 
  ylab( "p-value" ) 
@
  \caption{Difference in variability of \mb vs. multi-atlas segmentation accuracy. 
  In these plots we consider the segmentation accuracy within each subject across
  the 10 rounds of validation per parameter setting. Standard deviation and 
  variance is computed per subject for \mb and multi-atlas, and the distribution
  of these test statistics is compared with a t-test. The p-value of this test, 
  is shown on the y-axis (scaled logarithmically).}
  \label{ADNI1-xval-variability}
\end{figure}

@@

\begin{figure}
<<ADN1-xval-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep'>>=

bland_altman_plot <- function(data, x, y) {
  data$diff = data[x] - data[y] 
  return(ggplot(data, aes(x= x, y = diff)) + 
    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
               linetype=2, color='brown') + 
    geom_point())
}
a_20_tmpls = subset(all_data, 
                      reg_method   == "ANTS" &
                      method.mb    == "Majority Vote" &
                      templates.mb == 20 &
                      atlases      %in% seq(1,9,by =2) &
                      volume.ma     < 10000)
total_HC_volume = ddply(a_20_tmpls, c("subject", "label", "atlases"), function (df) {
  data.frame(
    ma = sum(df$volume.ma),
    mb = sum(df$volume.mb)
  )
})
total_HC_volume$diff = total_HC_volume$ma - total_HC_volume$mb

@@

% \begin{itemize}
% \item  A2A shows that if atlas population strongly(?) represents subject set
%        variability, then free choice from atlas population will produce
%        improvements (we know this b/c of extensive validation trials). 
% \item  what about in the case where atlas population doesn???t strongly
%        represent subject set variability (e.g. a priori atlas set)?  then, we
%        can use atlas selection to refine atlas set? 
% \end{itemize}
% 
% \todo{Kappa against our manual rater is low... explain that}

<<ADNI-src-volumes-prep,echo=F,cache=T>>=
#bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
#bl3T  = read.csv("data/ADNI_baseline_volumes/ADNI1_Baseline_3T_11_15_2012.csv")
#yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
maper = read.csv("data/ADNI_baseline_volumes/MAPER_volumes.csv")
snt   = read.csv("data/ADNI_baseline_volumes/UCSFSNTVOL.csv")
fs    = read.csv("data/ADNI_baseline_volumes/UCSFFSX_02_15_12.csv")
fsl   = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_FSLFIRST_volumes.csv", sep=';')
qc    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_QC_Park.csv", sep="\t")
# recode VISCODE
#viscode_levels = list("bl"=1,"bl"=2, "m06"=3,"m12"=4)
#yr1$VISCODE <- as.factor(yr1$Visit)
#levels(yr1$VISCODE) <-viscode_levels
#bl1.5T$Scaled_2 <- grepl("Scaled_2", yr1$Description)

#subjects = subset(yr1,Visit ==1)
subjects = bl
subjects$VISCODE = 'bl'

@@

# A table showing the number of images (and failures) from each package
#           Total  MAGeT  FSL  FS   MAPER
# Images    
# Failures
# 
# Total number complete images =

@@

attach(all_qc)
totals = data.frame(RID = RID, VISCODE = VISCODE, DX = factor(DX.bl),
                  MAGeT = MAGeT_L + MAGeT_R, 
                  MAPER = MAPER_L + MAPER_R, 
                  SNT   = SNT_L + SNT_R, 
                  #UCSD  = UCSD_L + UCSD_R,
                  FS    = FS_L + FS_R, 
                  FSL   = FSL_L + FSL_R)
detach(all_qc)
totals.complete = na.omit(totals)
@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1 Screening Dataset Package Totals & Failures                           %
%                        
%
\begin{figure}
<<ADNI1-scr-package-totals, echo=F, results="asis",cache=FALSE>>=
latex(package_totals, file="", size="scriptsize", 
    caption="ADNI1 1.5T Complete 1yr Dataset Segmentations",
    title="",
    label="tab:ADN1-1.5T-Complete-1yr-Dataset-Segmentations")
@
  \caption{Pass/fail quality control indicators were supplied with the FreeSurfer
  volumes downloaded from the ADNI website (we used the temporal lobe quality
  control indicator, TEMPQC). One of the authors (MP) performed visual 
  quality inspection for MAGeT and FSL segmentations.  
  }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T>>=
qplot(DX,value,data=melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT")),
      colour=variable,geom="boxplot") + 
      xlab("Diagnosis") + ylab("Total Hippocampal Volume (mm3)")
@
  \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), , FSL, 
  MAPER, \mb (MAGeT) and manual (SNT).}
  \label{ADNI-scr-volumes-boxplot}
\end{figure}

@@

latex(df, title="",
      n.cgroup=c(1,ncol(df)-1),cgroup=c("","Volume Correlation"),
      file="", size="scriptsize")
@

@@

\begin{figure}
<<FEP-volumes,cache=T>>=
SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
casted = dcast(SZ_volumes, Subject + Label ~ Method)
ggplot(data=casted, aes(x = Manual, y = MAGeT)) + 
  geom_smooth(method="lm") + 
  geom_point() + 
#   geom_text(aes(
#       label=paste("r = ", round(cor(Manual,MAGeT, use="complete.obs"), 2), 
#                   ", p <", round(as.numeric(t.test(Manual, MAGeT, paired=T)["p.value"]), 6))),
#       x=7000,y=10000, hjust=0,vjust=-0.5,size=5,data=data.frame()) +
   xlab("Total Hippocampal Volume (Manual Segmentation)") + 
   ylab("Total Hippocampal Volume (MAGeT Segmentation)")
)
@
  \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total
  HC volumes for \mb against manually rated Hippocampal volumes}
  \label{SZ_volumes}
\end{figure}

<<ADN1-FEP-Bland-Atlman, cache=TRUE, dependson='FEP-volumes'>>=
bland_altman_plot(casted$Manual, casted$MAGeT) + 
  xlab('Mean manual and MAGeT Brain total Hippocampus volume') + 
  ylab('Manual - MAGeT Brain volume') 
@
  \caption{Bland-Altman plot comparing manual and \mb hippocampal volumes 
  when using five Winterburn atlases, and a 21 image template library.}
  \label{}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Sweave/Knitr init
<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
opts_chunk$set(echo = FALSE, 
               message = FALSE, 
               warning = FALSE, 
               fig.align = 'center', 
               tidy = FALSE, 
               comment = NA, 
               cache = TRUE,
               fig.width = 6, 
               fig.height = 6)
library(ggplot2)
library(Hmisc)
library(reshape2)
library(ADNIMERGE)
library(plyr)
options(digits=3)
theme_set(theme_bw())

@@

\newcommand{\mb}{MAGeT-Brain }
\newcommand{\ants}{ANTS }
\newcommand{\animal}{ANIMAL }

@@

\subsubsection{ADNI1 1.5T Complete 1Yr Dataset}
Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
standardized dataset available from ADNI \footnote{
\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
\citep{Wyman2012}. This image collection contains uniformly preprocessed images
which have been designated to be the "best" after quality control.  All images
were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
Systems or Siemens Medical Solutions) at multiple sites using the protocol
described in \citep{Jack2008a}.  Representative 1.5T imaging parameters were TR
= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
x 240mm, a 192 x 192 x 166 matrix (x, y, and z directions) yielding a voxel
resolution of 1.25 x 1.25 x 1.2 $mm^3$. Clinical and demographic data are shown
in table \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Demographics}.

@@

latex(tab, file="", size="scriptsize", 
    caption="ADNI1 1.5T Complete 1Yr dataset demographics",
    label="tab:ADN1-1.5T-Complete-1Yr-Dataset-Demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Scanning was carried out at the Montreal Neurological Institute on a 1.5-T
Siemens whole body MRI system.  Structural T1 volumes were acquired for each
participant using a three-dimensional (3D) gradient echo pulse sequence with
sagittal volume excitation (repetition time=22ms, echo time=9.2ms, flip
angle=$30{\circ}$, 180 1mm contiguous sagittal slices). The rectangular field-of-view
for the images was 256mm (SI)×204mm (AP).  Subject demographics are shown in
table \ref{tab:SZFEP-Demographics}.

@@

<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
               data=fep_sz_demographics, method = "reverse", test = FALSE)
latex(tab, file="", ctable=FALSE,
      caption="Schizophrenia First Episode Patient Demographics",
      title="tab:SZFEP-Demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The \animal algorithm carries out image registration is in two phases. In the
first, a 12-parameter linear transformation (3 translations, rotations, scales,
shears) is estimated between images using an algorithm that maximizes the
correlation between blurred MR intensities and gradient magnitude over the whole
brain \citep{Collins}. In the second phase, nonlinear registration is completed
using the \animal algorithm \citep{Collins1995}: an iterative procedure that
estimates a 3D deformation field between two MR images. At first, large
deformations are estimated using blurred version of the input data. These larger
deformations are then input to subsequent steps where the fit is refined by
estimating smaller deformations on data blurred with a Gaussian kernel with a
smaller FWHM. The final transformation is a set of local translations defined on
a bed of equally spaced nodes that were estimated through the optimization of
the correlation coefficient.  For the purposes of this work we used the
regularization parameters optimized in Robbins et al. \citep{Robbins2004},
displayed in table \ref{tab:ANIMAL-params}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ANIMAL parameters                                                           %
%                                                                             %
\begin{table}[!tbp]
\begin{center}
\caption{\animal  registration parameters}
\begin{tabular}{l | c c c}
\hline 
Parameters        & Stage 1 & Stage 2 & Stage 3  \tabularnewline
\hline 
Model Blur (FWHM) &   8     &    8    &   4      \tabularnewline
Input Blur (FWHM) &   8     &    8    &   4      \tabularnewline
Iterations        &   30    &    30   &   10     \tabularnewline
Step              &   8x8x8 &  4x4x4  &   2x2x2  \tabularnewline
Sub-Lattice       &   6     &    6    &   6      \tabularnewline
Lattice Diameter  &24x24x24 &12x12x12 & 6x6x6    \tabularnewline
\hline
\end{tabular}
\end{center}
\label{tab:ANIMAL-params}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

{\bf Dataset evaluated.} 69 1.5T images were arbitrarily selected from the {\em
ADNI1:Complete 1Yr 1.5T} standardized dataset. 23 subjects were chosen from each
disease category (cognitively normal (CN), MCI and AD). Demographics for this
subset are shown in Table \ref{tab:ADNI1-xvalidation-demographics}.

@@

latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption="ADNI-1 cross-validation subset demographics", 
      title="tab:ADNI1-xvalidation-demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

{\bf Atlas and template libary.} The Winterburn atlases were used as the atlas
library.  The template library consisted of 21 randomly selected images from the
ADNI1 data dataset (7 healthy, MCI and AD subjects). Clinical and demographic
data for the template library subjects are shown in Table
\ref{tab:ADNI1-scr-tmpllib-demographics}.

@@

latex(tab, landscape=FALSE, ctable=FALSE, file="", size="scriptsize",
      caption="ADNI1 Complete 1Yr Template Library demographics", 
      title="tab:ADNI1-scr-tmpllib-demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}
<<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ggplot(all_data_mean, 
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + 
  facet_grid(reg_method~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1)) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean DSC" ) + 
  opts(legend.direction = "horizontal", legend.position = "bottom")
@
  \label{fig:ADNI1-xval-k_mean}
  \caption{
  {\bf Mean DSC by atlas and template library size, registration algorithm,
  and label fusion method.} The mean DSC score is calculated over all subjects 
  from the ADNI1 Cross-Validation dataset for each parameter setting.}
\end{figure}

@@

\begin{figure}
<<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ggplot(subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &
                templates.mb %in% seq(1,20,2)), 
       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
  #facet_grid(reg_method~method.y) + 
  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
  stat_summary(fun.y=mean, geom="point", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  stat_summary(fun.y=mean, geom="line", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  #stat_smooth(method=lm, formula=y~log(x)) + 
  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(legend=FALSE) + 
  scale_linetype_discrete(legend=FALSE) + 
  xlab( "Number of Templates" ) + 
  ylab( "Increase in mean DSC between MAGeT brain and multiatlas" ) +
  opts(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{The difference in mean DSC between \mb and multi-atlas segmentations for 
  a range of parameter settings.}
  \label{ADNI1-xval-k-diff}
\end{figure}

@@

ggplot(stats.tests, 
  aes(x=templates.mb, y=p.value, colour=as.factor(atlases), 
      size = abs(statistic))) + 
  facet_grid( . ~ stat) + 
  geom_line(aes(size=0.1,area=0.1)) + 
  geom_point(aes(shape = factor(direction))) + 
  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
  scale_shape_discrete("t-statistic sign", labels=c("positive","negative")) + 
  scale_size_area(name="t-statistic magnitude") + 
  scale_x_continuous(breaks=seq(1,20,by=2)) + 
  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
  xlab( "Number of Templates" ) + 
  ylab( "p-value" ) +
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(guide='none') + 
  scale_linetype_discrete(guide='none') + 
  scale_size_continuous(guide='none') + 
  opts(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{{\bf Difference in variability of \mb vs. multi-atlas segmentation 
  accuracy.}
  In these plots we consider the segmentation accuracy within each subject across
  the 10 rounds of validation per parameter setting. Standard deviation and 
  variance is computed per subject for \mb and multi-atlas, and the distribution
  of these test statistics is compared with a t-test. The p-value of this test, 
  is shown on the y-axis (scaled logarithmically).}
  \label{ADNI1-xval-variability}
\end{figure}

@@

\begin{figure}
<<ADN1-xval-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=8, fig.height=4>>=
a_20_tmpls = subset(all_data, 
                      reg_method   == "ANTS" &
                      method.mb    == "Majority Vote" &
                      templates.mb == 20 &
                      atlases      %in% seq(1,9,by =2) &
                      volume.ma     < 10000)
total_HC_volume = ddply(a_20_tmpls, c("subject", "label", "atlases"), function (df) {
  data.frame(
    ma = sum(df$volume.ma),
    mb = sum(df$volume.mb)
  )
})
total_HC_volume$diff = total_HC_volume$ma - total_HC_volume$mb

@@

<<ADNI-src-volumes-prep,echo=F,cache=T>>=
bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
maper = read.csv("data/ADNI_baseline_volumes/MAPER_volumes.csv")
snt   = read.csv("data/ADNI_baseline_volumes/UCSFSNTVOL.csv")
fs    = read.csv("data/ADNI_baseline_volumes/UCSFFSX_02_15_12.csv")
fsl   = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_FSLFIRST_volumes.csv", sep=';')
qc    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_QC_Park.csv", sep="\t")

subjects = bl
subjects$VISCODE = 'bl' # to match up with adnimerge

@@

@@

attach(all_qc)
totals = data.frame(RID = RID, VISCODE = VISCODE, DX = factor(DX.bl),
                  MAGeT = MAGeT_L + MAGeT_R, 
                  MAPER = MAPER_L + MAPER_R, 
                  SNT   = SNT_L + SNT_R, 
                  FS    = FS_L + FS_R, 
                  FSL   = FSL_L + FSL_R)
detach(all_qc)
totals.complete = na.omit(totals)
@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1 1yr Complete Dataset Package Totals & Failures                        %
%                        
%
<<ADNI1-scr-package-totals, echo=F, results="asis",cache=FALSE>>=
caption = paste(
  "Pass/fail quality control indicators were supplied with the FreeSurfer",
  "volumes downloaded from the ADNI website (we used the temporal lobe quality",
  "control indicator, TEMPQC). One of the authors (MP) performed visual",
  "quality inspection for MAGeT and FSL segmentations.", sep=" ")
latex(package_totals, file="", size="scriptsize", 
    caption = caption,
    title = "",
    label="tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Comparing total hippocampal volume across segmentation methods, we find close
agreement between manual volumes and MAPER and MAGeT volumes across disease 
categories (Figure \ref{ADNI-scr-volumes-boxplot}). FSL and FreeSurfer both 
produce volumes which are demonstratively larger than manually segmented volumes.

\begin{figure}[h]
<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
melted.by.snt = melt(totals.complete, measure.vars=c("MAGeT","FSL", "FS", "MAPER"))
snt.tests <- ddply(melted.by.snt, c("variable", "DX"), function(df) {
  data.frame( c = cor(df$SNT, df$value) ) 
})

qplot(DX,value,data=melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT")),
      colour=variable,geom="boxplot") + 
      xlab("Diagnosis") + ylab("Total Hippocampal Volume (mm3)")
@
  \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), FSL, 
  MAPER, \mb (MAGeT) and manual (SNT).}
  \label{ADNI-scr-volumes-boxplot}
\end{figure}

@@

#latex(df, title="",
#      n.cgroup=c(1,ncol(df)-1),cgroup=c("","Volume Correlation"),
#      file="", size="scriptsize")
@

@@

\begin{figure}
<<FEP-volumes,cache=T>>=
SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
casted = dcast(SZ_volumes, Subject + Label ~ Method)
ggplot(data=casted, aes(x = Manual, y = MAGeT)) + 
  geom_smooth(method="lm") + 
  geom_point() + 
#   geom_text(aes(
#       label=paste("r = ", round(cor(Manual,MAGeT, use="complete.obs"), 2), 
#                   ", p <", round(as.numeric(t.test(Manual, MAGeT, paired=T)["p.value"]), 6))),
#       x=7000,y=10000, hjust=0,vjust=-0.5,size=5,data=data.frame()) +
   xlab("Total Hippocampal Volume (Manual Segmentation)") + 
   ylab("Total Hippocampal Volume (MAGeT Segmentation)")
@
  \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total
  HC volumes for \mb against manually rated Hippocampal volumes}
  \label{SZ_volumes}
\end{figure}

\begin{figure}
<<ADN1-FEP-Bland-Atlman, cache=TRUE, dependson='FEP-volumes'>>=
bland_altman_plot(casted$Manual, casted$MAGeT) + 
  xlab('Mean manual and MAGeT Brain total Hippocampus volume') + 
  ylab('Manual - MAGeT Brain volume') 
@
  \caption{Bland-Altman plot comparing manual and \mb hippocampal volumes 
  when using five Winterburn atlases, and a 21 image template library.}
  \label{}
\end{figure}
